---
title: "Lab 5 assignment - Point pattern analysis"
output:
  html_document:
    df_print: paged
---

```{r}
require(spatstat)
require(tidyverse)
require(terra)
```


# Challenge 1 (3 points)

Create a simulation window on a landscape stretching from 0 to 100 units in both the x and y directions. Now simulate the distribution of an organism on this landscape using a Poisson process with a lambda value of 0.01. Plot this organism's distribution. How many organisms are there on your landscape? What is the expected number of organisms that you should find on this landscape based on your simulation parameters? Why?

```{r}
#Put your code here.

#Creating window
pppWindow=owin(xrange=c(0,100),yrange=c(0,100))

#Creating simulated distribution of organism on the created window
set.seed(1)
simDist = rpoispp(lambda = 0.01, nsim=1, win=pppWindow)
plot(simDist)
#Should be 100 
#We have 93
```

The expected number of points would be 100 points.The lamda argument specifies the number of points per unit area of the window. Our window is 100*100, which we then multiply by the lamda value of 0.01 which equals 100 points. Our specific simulation had 93 points in the window which is slightly lower than expected. 

$\color{red}{\text{Perfect. +3}}$


# Challenge 2 (3 points)

Verify that there no linear or quadratic pattern in the distribution of your organisms using fitted Poisson process models. Show the outcome of this analysis and explain your conclusion.

```{r}

ppInt = ppm(simDist, ~1)
ppXy = ppm(simDist, ~ x + y)
ppXy2 = ppm(simDist, ~polynom (x, y, 2))

data.frame(model = c('int', 'xy', 'xy2'),
           AIC = c(AIC(ppInt), AIC(ppXy), AIC(ppXy2)))
#Homogenous model has the most support (lowest AIC score)

#Plots just for fun
plot(predict(ppXy2, type='trend'))
plot(simDist, add=T)

plot(predict(ppXy, type='trend'))
plot(simDist, add=T)

plot(predict(ppInt, type='trend'))
plot(simDist, add=T)
```

I fitted a homogenous, linear, and quadratic model to the simulated point distribution. THe homogenous model had the most support (the lowest AIC) indicating that there is likely no significant evidence of linear or quadratic patterning in the distribution of our simulated organisms. 

$\color{red}{\text{Nice. +3}}$


# Challenge 3 (14 points)

I have simulated the distributions of 2 critters and provided their x-y coordinates. One of them is a non-territorial predator. The second is a critter that is preyed upon by the predator and exhibits a fear response to the predator's presence. Use the tools you learned in the lab to explore the distributions of these two critters. Identify which one is the predator and which is the prey, and provide code and a thorough justification supporting your conclusion.


```{r}
#Use this code chunk to get you started.

predPreyWin = owin(xrange = c(0, 100), yrange = c(0, 100))

critter1 = read.csv('https://raw.githubusercontent.com/ValenteJJ/SpatialEcology/main/Week5/species1.csv') %>% 
  mutate(org = 'critter1')

critter2 = read.csv('https://raw.githubusercontent.com/ValenteJJ/SpatialEcology/main/Week5/species2.csv') %>% 
  mutate(org = 'critter2')


#Creating point patterns for the 2 critters
pppCritter1=ppp(x=critter1$x,y=critter1$y,window=predPreyWin)
plot(pppCritter1)

pppCritter2=ppp(x=critter2$x,y=critter2$y,window=predPreyWin)
plot(pppCritter2)

#Plotting density 
plot(pppCritter1)
plot(density(pppCritter1, 1))

plot(pppCritter2)
plot(density(pppCritter2,1))

#Summary Stats
summary(pppCritter1)
summary(pppCritter2)

#quadrat counts
q_critter1 = quadratcount(pppCritter1, nx=4, ny=4)
plot(pppCritter1)
plot(q_critter1, add=T)

q_critter2 = quadratcount(pppCritter2,nx=4, ny=4)
plot(pppCritter2)
plot(q_critter2, add=T)

#Ripleys K and L no edge corrections
  #Ripleys K
  critter1_kNone = Kest(pppCritter1, correction='none')
  plot(critter1_kNone)
  critter2_kNone = Kest(pppCritter2, correction='none')
  plot(critter2_kNone)

  
  #Ripleys L
  critter1_lNone = Lest(pppCritter1, correction='none')
  critter2_lNone = Lest(pppCritter2, correction='none')
  
  
  #Increases with radius
  plot(critter1_lNone)
  plot(critter2_lNone)
  
  
  #Subtract r
  plot(critter1_lNone, .-r~r)
  plot(critter2_lNone, .-r~r)

  
  #Isotropic edge correction for critters
  critter1_lIso = Lest(pppCritter1, correction = 'isotropic')
  plot(critter1_lIso, . -r~r)
  
  critter2_lIso = Lest(pppCritter2, correction = 'isotropic')
  plot(critter2_lIso, . -r~r)


  #Observed pattern statistically different from CSR?
  critter1_lCsr = envelope(pppCritter1, Lest, nsim=99, rank=1, correction='isotropic',
                            global=F)
  plot(critter1_lCsr, .-r~r, shade=c('hi', 'lo'), legend=F)

  
  #Observed pattern appears to not be statistically different from CSR
  critter2_lCsr = envelope(pppCritter2, Lest, nsim=99, rank=1, correction='isotropic',
                            global=F)
  plot(critter2_lCsr, .-r~r, shade=c('hi', 'lo'), legend=F)
  
  
  #Pair correlation function
  critter1_pTrans = pcf(pppCritter1, correction='isotropic')
  plot(critter1_pTrans)
  critter2_pTrans = pcf(pppCritter2, correction='isotropic')
  plot(critter2_pTrans)
  
  critter1_pEnv = envelope(pppCritter1, pcf, nsim=99, rank=1, correction='isotropic',
                           global=F)
  plot(critter1_pEnv, shade=c('hi', 'lo'), legend=F)
  critter2_pEnv = envelope(pppCritter2, pcf, nsim=99, rank=1, correction='isotropic',
                           global=F)
  plot(critter2_pEnv, shade=c('hi', 'lo'), legend=F)  
  
  
  #G Function (nearest neighbors)
  critter1_gTrans = Gest(pppCritter1, correction='rs')
  plot(critter1_gTrans, legend=F)
  critter2_gTrans = Gest(pppCritter2, correction='rs')
  plot(critter2_gTrans, legend=F)
  
  critter1_gEnv = envelope(pppCritter1, Gest, nsim=99, rank=1, correction='rs',
                           global=F)
  plot(critter1_gEnv, shade=c('hi', 'lo'), legend=F)
  critter2_gEnv = envelope(pppCritter2, Gest, nsim=99, rank=1, correction='rs',
                           global=F)
  plot(critter2_gEnv, shade=c('hi', 'lo'), legend=F)
  
  #Lcross
  
  #Creating new dataframe combining the 2 critters
  critters=rbind(critter1,critter2)
  critters$org=as.factor(critters$org)
  
  #New PPP object of the two 
  pppCritters = ppp(critters$x, critters$y, window = predPreyWin,marks=critters$org)
  plot(pppCritters)
  
  
  critters_lMulti = envelope(pppCritters, Lcross, nsim=99, rank=1, global=F,
                             i='critter1',
                    simulate=expression(rlabel(pppCritters)))
  plot(critters_lMulti, .-r~r, legend=F)
```
Ripley's K didn't really show us too much. Both critters roughly follow what was expected. Ripley's L (with edge correction) however, paints a much better picture of the species. Critter 1 seems especially different from what would be expected from random generation. Furthermore, it is almost completely outside the envelope we created, meaning it is significantly different from the expected. Critter 2, however, was completely inside the envelope for Ripley's L, meaning we cannot necessarily differentiate that Critter 2 isn't significantly different from random generation. G values didn't really tell us anything at all. When running the lcross function, we found that how far critter 2 was from critter 1 was significantly different than would be predicted if it were randomly generated. 

Given what I found out through this analysis, I would guess that Critter 1 is the prey species. The fact that critter 1s points were distributed significantly differently from what would be randomly expected whereas critter 2 was not is a important detail. I would expect a prey species to be congregated in small areas (clumped) as they get chased by predators and thus not have a random distribution across the landscape which is consistent with what I found for Critter 1.  Critter 2, had a roughly random distribution which is what I would probably expect from a predator that isn't getting chased down into small areas. A predator can roam more freely and thus more random. 


$\color{red}{\text{Excellent work, Marcus. +14}}$



