---
title: "Spatial Ecology Final Project"
author: "Marcus Prull"
date: "4/11/2024"
output: html_document
---

First I will start by loading required packages.
```{r setup, include=FALSE}
require(sf)
require(raster)
require(terra)
require(gdistance)
require(dplyr)
require(lubridate)
require(ggplot2)
```
Here is where I make the transition/cost layer from my already created Neely Henry Shapefile.
```{r}
  #Reading in Neely ShapeFile
  Neely = st_read("/vsicurl/https://github.com/PrullMarcus/SpatialEcology/raw/main/FinalProject/NeelySHP_NAD_1983_2011.shp")
  plot(Neely$geometry)

  #First need to convert Neely shapefile into "cost" raster
  NeelyVec=vect(Neely)
  NeelyR=rast(NeelyVec, res=0.00003)#00005 = 5.5 m cell. Running about a 4.4m grid cell (0.00004), 0.00001 would be      1.1m cells, would like to be able to get down to maybe 2 m 
  NeelyRast=rasterize(NeelyVec,NeelyR)#value of 1 is Neely, value of NA is land
  plot(NeelyRast)#looks correct

  #Creating transition layer of Neely Thing
  NeelyRast = raster(NeelyRast)#Converts spatraster to "raster" that can be used by transition function
  NeelyTrans = transition(NeelyRast, transitionFunction = mean, 8)
  
  #Correcting for corners
  NeelyTrans = geoCorrection(NeelyTrans, type='c', multpl=F)

  #Visualizing Cost/resistance raster
  NeelyTransRast = raster(NeelyTrans)
  plot(NeelyTransRast)
```
Reading in updated Dispersal data with GroupMort Column to sort by Alive vs Dead fish.
```{r}
Dispersal_raw = "https://github.com/PrullMarcus/SpatialEcology/raw/main/FinalProject/DispersalFishONLY.csv"
Dispersal_raw = read.csv(Dispersal_raw)
Dispersal_raw = subset(Dispersal_raw,GroupMovement=="A")#Gets rid of dead fish for my analysis

#Now I need to convert my lat longs into spatial points objects with the correct projection
Dispersal_sf = st_as_sf(Dispersal_raw, coords = c("Long","Lat"))

#Defining its coordinate system
CoordSys = st_crs(Neely)#Retrieving CS from Neely Shapefile (Should be the same as these points)
Dispersal_sf = st_set_crs(Dispersal_sf, CoordSys)#Setting it to the right coord system.

#Plotting to make sure it shows up in the right spot
plot(Neely$geometry)
plot(Dispersal_sf, add = T)#Looks good to me
```

Filtering Data, and then loading in boat launch locations
```{r}
#All files are in NAD83(2011) 

#Filtering Data Appropriately
DispersalFilt = Dispersal_sf %>% group_by(RewardTag1) %>% filter(n()>1)#Filtering to only fish that have more than 1 observation

#Checking to make sure no points fall inside polygon
as.integer(st_intersects(DispersalFilt,Neely))#all values gave me a 1 (all inside polygon)

#Making a proper "Date" column for use later
Dates = make_date(year = DispersalFilt$Year, month = DispersalFilt$Month, day = DispersalFilt$Day)
DispersalFilt = cbind(Dates,DispersalFilt)
colnames(DispersalFilt)[1]<-"Date"    #Renaming date column to "Date"

#Reading in main boat launch location (Coosa Landing aka "CL")
CL = st_read("/vsicurl/https://github.com/PrullMarcus/SpatialEcology/raw/main/FinalProject/CL.shp")

#Sorting my release fish to only those that were released at "CL"
CL_Fish = subset(DispersalFilt,Ramp=="CL")
```

Here I will actually calculate least cost distance for individuals dispersing from tournament boat ramps. The least cost distance will be indicative of the individuals proximity from the ramp. 
```{r}
#### COOSA LANDING PROXIMITY ####
CL_prox_m = numeric(nrow(CL_Fish))
length(CL_prox_m)#Checks out

for (i in 1:nrow(CL_Fish)){
  
  CL_prox_m[i] = gdistance::costDistance(NeelyTrans, fromCoords = sf::st_coordinates(CL), 
                            toCoords = sf::st_coordinates(CL_Fish[i,]))
      }
    
    
CL_disp = cbind(CL_prox_m,CL_Fish)


tag_numbers = unique(CL_Fish$RewardTag1)#unique tag numbers
n_fish = length(tag_numbers)#number of fish total
outputDaysSinceRel=numeric()


for(i in 1:n_fish){
  
  new_tag = tag_numbers[i]
  new_dat = subset(CL_Fish,RewardTag1 == new_tag)
  tmax = nrow(new_dat)
  days = numeric(tmax)
  
  for(t in 2:tmax){
    
    days[t] = as.numeric(new_dat$Date[t]-new_dat$Date[1])
    
  }
  outputDaysSinceRel = append(outputDaysSinceRel,days)
}

CL_disp=cbind(outputDaysSinceRel,CL_disp)
colnames(CL_disp)[1]="DaysSinceRel"

CL_disp$WeeksSinceRel=CL_disp$DaysSinceRel/7
```

Now I will add the speed element in m/w. This will look at how their movement speeds change over time after they are released from the boat launch. To do this, I will calculate least cost distance between successive locations for each individual fish. Then, I will need to slice the first row of data and calculate days post release. For each individual. I should just be able to do this for all fish at once, instead of separating by boat launch like I did in the previous part. 

```{r}
#Creating new objects to be used in loop
tag_numbers = unique(CL_Fish$RewardTag1)#unique tag numbers
n_fish = length(tag_numbers)#number of fish total

#Creating vectors for output to be stored in 
outputMovement = numeric()
outputDaysBetweenLocs = numeric()

#Outside loop that selects tag-number 
for(i in 1:n_fish){
  
  new_tag = tag_numbers[i]
  new_dat = subset(CL_Fish,RewardTag1 == new_tag)
  tmax = nrow(new_dat)
  distbetweenlocs = numeric(tmax)
  daysbetweenlocs = numeric(tmax)
  
#Inside loops over time for a given tag-number  
  for(t in 2:tmax){
    
    dist_m = gdistance::costDistance(NeelyTrans, fromCoords = sf::st_coordinates(new_dat[t-1,]), 
                                                 toCoords = sf:: st_coordinates(new_dat[t,]))
    distbetweenlocs[t] = dist_m
    daysbetweenlocs[t] = as.numeric(new_dat$Date[t]-new_dat$Date[t-1])
    
  }
  outputMovement = append(outputMovement,distbetweenlocs)
  outputDaysBetweenLocs = append(outputDaysBetweenLocs,daysbetweenlocs)
}
outputMovement
outputDaysBetweenLocs

#Adding distances to new "CL_Movement" dataframe
CL_Final = cbind(outputMovement,CL_disp)
colnames(CL_Final)[1]<-"MetersFromLastLocation"

#Adding days post release to the "DispersalMovement" dataframe
CL_Final = cbind(outputDaysBetweenLocs,CL_Final)
colnames(CL_Final)[1]<-"DaysFromLastLoc"

#Adding weeks between successive locations as well as their m/week movement speeds
CL_Final$WeeksFromLastLoc = CL_Final$DaysFromLastLoc/7
CL_Final$MetersPerWeek = CL_Final$MetersFromLastLocation/CL_Final$WeeksFromLastLoc

#Converting to Dataframe to be written into a .csv
CL_Dispersal=as.data.frame(CL_Final)

#Simplifying dataframe down to exclude geometry (the geometry messed with the columns)
CL_Dispersal_FINAL = select(CL_Dispersal,c(DaysFromLastLoc,MetersFromLastLocation,DaysSinceRel,CL_prox_m,Date,TagSite,Species,TL,RewardTag1,Combined,Radio.Tag.Frequency,Ramp,WeeksSinceRel,WeeksFromLastLoc,MetersPerWeek,A.D,GroupMovement))

#Writing csv to be able to make graphs of calculated data at home. My computer was too feeble to be able to run the last cost analysis stuff without blowing up so I had to do it on a desktop at work. 
#write.csv(CL_Dispersal_FINAL,"C:/Users/mrp0099/Desktop/MarcusStuff/SpatialEcology/CL_FINAL2.csv")
```


Here will be the plots for Dispersal Stuff. (Prox to ramp over time)(Movement speeds over time)(Prop of individuals w/in a specific prox)
```{r}







```

Loading in control data and creating control points object.
```{r}
Control_raw = "https://github.com/PrullMarcus/SpatialEcology/raw/main/FinalProject/Control%20Fish%202023%2C%2077%20original.csv"
Control_raw = read.csv(Control_raw)
Control_raw = subset(Control_raw,GroupMovement=="A")#Gets rid of dead fish for my analysis

#Now I need to convert my lat longs into spatial points objects with the correct projection
Control_sf = st_as_sf(Control_raw, coords = c("Long","Lat"))

#Defining its coordinate system
CoordSys = st_crs(Neely)#Retrieving CS from Neely Shapefile (Should be the same as these points)
Control_sf = st_set_crs(Control_sf, CoordSys)#Setting it to the right coord system.

#Plotting to make sure it shows up in the right spot
plot(Neely$geometry)
plot(Control_sf, add = T)#Looks good to me
```

Here I will be doing similar things for the control fish. However, I am more focused on how their movement speeds change over time (looking for seasonal trends). 
```{r}
ControlFilt = Control_sf %>% group_by(RewardTag1) %>% filter(n()>2) %>% slice(2:n())#Filtering to only fish that have 3 or more observations and then removing the first row from each fish that qualifies (the first row is its initial tagging location)

#Converting year month day columns into date column to be used for elapsed time (get movement/time rates)
Dates = make_date(year = ControlFilt$Year, month = ControlFilt$Month, day = ControlFilt$Day)
ControlFilt = cbind(Dates,ControlFilt)
colnames(ControlFilt)[1]<-"Date"    #Renaming date column to "Date"

#Creating objects to be used by loop
tag_numbers = unique(ControlFilt$RewardTag1)#unique tag numbers
n_fish = length(tag_numbers)#number of fish total

#Creating output vectors
outputControlMovement=numeric()
outputDaysBetweenLocs=numeric()

#Outside loop that selects tag-number 
for(i in 1:n_fish){
  
  new_tag = tag_numbers[i]
  new_dat = subset(ControlFilt,RewardTag1 == new_tag)
  tmax = nrow(new_dat)
  ControlDist = numeric(tmax)
  days = numeric(tmax)
  
  for(t in 2:tmax){
    
    dist_m = gdistance::costDistance(NeelyTrans, fromCoords = sf::st_coordinates(new_dat[t-1,]), 
                                                 toCoords = sf:: st_coordinates(new_dat[t,]))
    ControlDist[t] = dist_m
    days[t] = as.numeric(new_dat$Date[t]-new_dat$Date[t-1])
    
  }
  outputControlMovement = append(outputControlMovement,ControlDist)
  outputDaysBetweenLocs = append(outputDaysBetweenLocs,days)
}
outputControlMovement
outputDaysBetweenLocs

ControlMovement = cbind(outputControlMovement,ControlFilt)
colnames(ControlMovement)[1]<-"MetersFromLastLocation"

ControlMovement = cbind(outputDaysBetweenLocs,ControlMovement)
colnames(ControlMovement)[1]<-"DaysSinceLastLoc"
names(ControlMovement)

length(unique(ControlMovement$RewardTag1))

#Removing the first location (no movement value here just 0s) should remove 80 rows of data thus bringing the total to 686 total
FinalControlMovement = ControlMovement %>% group_by(RewardTag1) %>% slice(2:n())

#Converting days between loc to weeks since last loc
FinalControlMovement$WeeksSinceLastLoc = FinalControlMovement$DaysSinceLastLoc/7

#Getting movement speeds (meters/week)
FinalControlMovement$MetersPerWeek = FinalControlMovement$MetersFromLastLocation/FinalControlMovement$WeeksSinceLastLoc

#Simplifying the data frame to be exported for use by me at home. (removing geometry column)
Control_FINAL = as.data.frame(FinalControlMovement)
Control_FINAL = select(Control_FINAL,c(DaysSinceLastLoc,MetersFromLastLocation,Date,TagSite,Species,TL,RewardTag1,Combined,Radio.Tag.Frequency,WeeksSinceLastLoc,MetersPerWeek,A.D,GroupMovement))

#CSV for me to make graphs at home later.
write.csv(Control_FINAL,"C:/Users/mrp0099/Desktop/MarcusStuff/SpatialEcology/Control_FINAL.csv")
```

```{r}

#Round to nearest week and then look at how average movement speed differs over time for dispersed individuals 
#Construct homeranges
#Step-selection model to assess efficiency of individuals dispersing. 

```

Here I will attempt to construct homeranges for fish in my system. (Kernel Density) 
```{r}



```
