---
title: "Spatial Ecology Final Project"
author: "Marcus Prull"
date: "4/11/2024"
output: html_document
---

First I will start by loading required packages.
```{r setup, include=FALSE}
require(sf)
require(raster)
require(terra)
require(gdistance)
require(dplyr)
require(lubridate)
require(ggplot2)
```
Here is where I make the transition/cost layer from my already created Neely Henry Shapefile.
```{r}
  #Reading in Neely ShapeFile
  Neely = st_read("/vsicurl/https://github.com/PrullMarcus/SpatialEcology/raw/main/FinalProject/NeelySHP_NAD_1983_2011.shp")
  plot(Neely$geometry)

  #First need to convert Neely shapefile into "cost" raster
  NeelyVec=vect(Neely)
  NeelyR=rast(NeelyVec, res=0.00004)#00005 = 5.5 m cell. Running about a 4.4m grid cell (0.00004), 0.00001 would be      1.1m cells, would like to be able to get down to maybe 2 m 
  NeelyRast=rasterize(NeelyVec,NeelyR)#value of 1 is Neely, value of NA is land
  plot(NeelyRast)#looks correct

  #Creating transition layer of Neely Thing
  NeelyRast = raster(NeelyRast)#Converts spatraster to "raster" that can be used by transition function
  NeelyTrans = transition(NeelyRast, transitionFunction = mean, 8)
  
  #Correcting for corners
  NeelyTrans = geoCorrection(NeelyTrans, type='c', multpl=F)

  #Visualizing Cost/resistance raster
  NeelyTransRast = raster(NeelyTrans)
  plot(NeelyTransRast)
```
Reading in updated Dispersal data with GroupMort Column to sort by Alive vs Dead fish.
```{r}
Dispersal_raw = "https://github.com/PrullMarcus/SpatialEcology/raw/main/FinalProject/DispersalFishONLY.csv"
Dispersal_raw = read.csv(Dispersal_raw)
Dispersal_raw = subset(Dispersal_raw,GroupMovement=="A")#Gets rid of dead fish for my analysis

#Now I need to convert my lat longs into spatial points objects with the correct projection
Dispersal_sf = st_as_sf(Dispersal_raw, coords = c("Long","Lat"))

#Defining its coordinate system
CoordSys = st_crs(Neely)#Retrieving CS from Neely Shapefile (Should be the same as these points)
Dispersal_sf = st_set_crs(Dispersal_sf, CoordSys)#Setting it to the right coord system.

#Plotting to make sure it shows up in the right spot
plot(Neely$geometry)
plot(Dispersal_sf, add = T)#Looks good to me
```

Filtering Data, and then loading in boat launch locations
```{r}
#All files are in NAD83(2011) 

#Filtering Data Appropriately
DispersalFilt = Dispersal_sf %>% group_by(RewardTag1) %>% filter(n()>1)#Filtering to only fish that have more than 1 observation

#Checking to make sure no points fall outside polygon
as.integer(st_intersects(DispersalFilt,Neely))#all values gave me a 1 (all inside polygon)

#Making a proper "Date" column for use later
Dates = make_date(year = DispersalFilt$Year, month = DispersalFilt$Month, day = DispersalFilt$Day)
DispersalFilt = cbind(Dates,DispersalFilt)
colnames(DispersalFilt)[1]<-"Date"    #Renaming date column to "Date"

#Confirming that points show up in shapefile
plot(NeelyRast)
plot(DispersalFilt,add=T)

extract(raster(NeelyRast),Dispersal_raw)

#Reading in boat launch/release location .shp files
CL = st_read("/vsicurl/https://github.com/PrullMarcus/SpatialEcology/raw/main/FinalProject/CL.shp")
SS = st_read("/vsicurl/https://github.com/PrullMarcus/SpatialEcology/raw/main/FinalProject/SS.shp")
RL = st_read("/vsicurl/https://github.com/PrullMarcus/SpatialEcology/raw/main/FinalProject/RL.shp")
CC = st_read("/vsicurl/https://github.com/PrullMarcus/SpatialEcology/raw/main/FinalProject/CC.shp")

#Sorting my fish by release location
CL_Fish = subset(Dispersal,Ramp=="CL")
SS_Fish = subset(Dispersal,Ramp=="SS")
RL_Fish = subset(Dispersal,Ramp=="RL")
CC_Fish = subset(Dispersal,Ramp=="CC")
```

Here I will actually calculate least cost distance for individuals dispersing from tournament boat ramps. The least cost distance will be indicative of the individuals proximity from the ramp. 
```{r}
#### COOSA LANDING PROXIMITY ####
CL_prox_m = numeric(nrow(CL_Fish))

for (i in 1:nrow(CL_Fish)){
  
  CL_prox_m[i] = gdistance::costDistance(NeelyTrans, fromCoords = sf::st_coordinates(CL), 
                            toCoords = sf::st_coordinates(CL_Fish[i,]))
      }
    
    
CL_distances = cbind(CL_prox_m,CL_Fish)
    


#### RAINBOW LANDING PROXIMITY ####
RL_prox_m = numeric(nrow(RL_Fish))

for (i in 1:nrow(RL_Fish)){
  
  RL_prox_m[i] = gdistance::costDistance(NeelyTrans, fromCoords = sf::st_coordinates(RL), 
                            toCoords = sf::st_coordinates(RL_Fish[i,]))
      }
    
    
RL_distances = cbind(RL_prox_m,RL_Fish)



#### SOUTHSIDE LANDING PROXIMITY ####
SS_prox_m = numeric(nrow(SS_Fish))

for (i in 1:nrow(SS_Fish)){
  
  SS_prox_m[i] = gdistance::costDistance(NeelyTrans, fromCoords = sf::st_coordinates(SS), 
                            toCoords = sf::st_coordinates(SS_Fish[i,]))
      }
    
    
SS_distances = cbind(SS_prox_m,SS_Fish)



#### CANOE CREEK PROXIMITY ####
CC_prox_m = numeric(nrow(CC_Fish))

for (i in 1:nrow(CC_Fish)){
  
  CC_prox_m[i] = gdistance::costDistance(NeelyTrans, fromCoords = sf::st_coordinates(CC), 
                            toCoords = sf::st_coordinates(CC_Fish[i,]))
      }
    
    
CC_distances = cbind(CC_prox_m,CC_Fish)

CL_proximities = CL_distances

write.csv(CL_proximities,"C:/Users/mrp0099/Desktop/MarcusStuff/SpatialEcology/CL_proximities.csv")
```

Now I will add the speed element in m/d or m/w. This will look at how their movement speeds/step length distances may change after they are released from a boat launch. To do this, I will calculate least cost distance between successive locations for each individual fish. Then, I will need to slice the first row of data and calculate days post release. For each individual. I should just be able to do this for all fish at once, instead of separating by boat launch like I did in the previous part. 

```{r}
#Creating new objects to be used in loop
tag_numbers = unique(DispersalFilt$RewardTag1)#unique tag numbers
n_fish = length(tag_numbers)#number of fish total

#Creating vectors for output to be stored in 
output = numeric()
outputDays = numeric()

#Outside loop that selects tag-number 
for(i in 1:n_fish){
  
  new_tag = tag_numbers[i]
  new_dat = subset(DispersalFilt,RewardTag1 == new_tag)
  tmax = nrow(new_dat)
  DispersalDist = numeric(tmax)
  days = numeric(tmax)
  
#Inside loops over time for a given tag-number  
  for(t in 2:tmax){
    
    dist_m = gdistance::costDistance(NeelyTrans, fromCoords = sf::st_coordinates(new_dat[t-1,]), 
                                                 toCoords = sf:: st_coordinates(new_dat[t,]))
    DispersalDist[t] = dist_m
    days[t] = as.numeric(new_dat$Date[t]-new_dat$Date[t-1])
    
  }
  output = append(output,DispersalDist)
  outputDays = append(outputDays,days)
}
output
outputDays

#Adding distances to new "DispersalMovement" dataframe
DispersalMovement = cbind(output,DispersalFilt)
colnames(DispersalMovement)[1]<-"MetersFromLastLocation"
names(DispersalMovement)

#Adding days post release to the "DispersalMovement" dataframe
DispersalMovement = cbind(outputDays,DispersalMovement)
colnames(DispersalMovement)[1]<-"DaysElapsed"
names(DispersalMovement)

FinalDispersalMovement = DispersalMovement %>% group_by(RewardTag1) %>% slice(2:n())

FinalDispersalMovement$WeeksElapsed = FinalDispersalMovement$DaysElapsed/7

FinalDispersalMovement$MetersPerDay = FinalDispersalMovement$MetersFromLastLocation/FinalDispersalMovement$DaysElapsed

FinalDispersalMovement$MetersPerWeek = FinalDispersalMovement$MetersFromLastLocation/FinalDispersalMovement$WeeksElapsed

Dispersal_Speeds=as.data.frame(FinalDispersalMovement)
```

Here I am doing the same thing for the control fish 
```{r}
Control = st_read("/vsicurl/https://github.com/PrullMarcus/SpatialEcology/raw/main/FinalProject/ControlFish_3_4_24.shp")

ControlFilt = Control %>% group_by(RewardTag1) %>% filter(n()>2)#Filtering to only fish that have more than 3 observations or more

#Converting year month day columns into date column to be used for elapsed time (get movement/time rates)
Dates = make_date(year = ControlFilt$Year, month = ControlFilt$Month, day = ControlFilt$Day)
ControlFilt = cbind(Dates,ControlFilt)
colnames(ControlFilt)[1]<-"Date"    #Renaming date column to "Date"

#Creating objects to be used by loop
tag_numbers = unique(ControlFilt$RewardTag1)#unique tag numbers
n_fish = length(tag_numbers)#number of fish total

#Creating output vectors
output2=numeric()
outputDays2=numeric()

#Outside loop that selects tag-number 
for(i in 1:n_fish){
  
  new_tag = tag_numbers[i]
  new_dat = subset(ControlFilt,RewardTag1 == new_tag)
  tmax = nrow(new_dat)
  ControlDist = numeric(tmax)
  days = numeric(tmax)
  
  for(t in 2:tmax){
    
    dist_m = gdistance::costDistance(NeelyTrans, fromCoords = sf::st_coordinates(new_dat[t-1,]), 
                                                 toCoords = sf:: st_coordinates(new_dat[t,]))
    ControlDist[t] = dist_m
    days[t] = as.numeric(new_dat$Date[t]-new_dat$Date[t-1])
    
  }
  output2 = append(output2,ControlDist)
  outputDays2 = append(outputDays2,days)
}
output2#spit out some infinities????? Neely shapefile not encompassing enough?????
outputDays2

n=nrow(ControlFilt)
ControlFilt2 = ControlFilt[1:918,]#6 rows at end of dataframe were empty

ControlMovement = cbind(output2,ControlFilt2)
colnames(ControlMovement)[1]<-"MetersFromLastLocation"
names(ControlMovement)

ControlMovement = cbind(outputDays2,ControlMovement)
colnames(ControlMovement)[1]<-"DaysElapsed"
names(ControlMovement)

FinalControlMovement = ControlMovement %>% group_by(RewardTag1) %>% slice(3:n())

FinalControlMovement$WeeksElapsed = FinalControlMovement$DaysElapsed/7

FinalControlMovement$MetersPerDay = FinalControlMovement$MetersFromLastLocation/FinalControlMovement$DaysElapsed

FinalControlMovement$MetersPerWeek = FinalControlMovement$MetersFromLastLocation/FinalControlMovement$WeeksElapsed

Control_Speeds = as.data.frame(FinalControlMovement)
write.csv(Control_Speeds,"C:/Users/mrp0099/Desktop/MarcusStuff/SpatialEcology/Control_Speeds.csv")
```

```{r}

#Round to nearest week and then look at how average movement speed differs over time for dispersed indivuals. 
#Compare this to control individuals for both species. 
#Construct homeranges
#Step-selection model to assess efficiency of individuals dispersing. 

```

Here I will attempt to construct homeranges for fish in my system. (Kernel Density) 
```{r}



```

Here I will create a plot of average proximity to the ramp over time
```{r}
#Average Dispersal Proxmity over time (of alive fish)
CL_Dispersals = read.csv("CL_proximities.csv")#fix this to github
CL_Disp_Alives = subset(CL_Dispersals,MortGroup == "A")
CL_Disp_Alives = CL_Disp_Alives %>% group_by(RewardTag1) %>% filter(n()>1)

Dates = make_date(year = CL_Disp_Alives$Year, month = CL_Disp_Alives$Month, day = CL_Disp_Alives$Day)
CL_Disp_Alives = cbind(Dates,CL_Disp_Alives)
colnames(CL_Disp_Alives)[1]<-"Date"    #Renaming date column to "Date"

tag_numbers = unique(CL_Disp_Alives$RewardTag1)#unique tag numbers
n_fish = length(tag_numbers)#number of fish total

outputDays2 = numeric()
for(i in 1:n_fish){
  
  new_tag = tag_numbers[i]
  new_dat = subset(CL_Disp_Alives,RewardTag1 == new_tag)
  tmax = nrow(new_dat)
  days = numeric(tmax)
  
  for(t in 2:tmax){
    
    days[t] = as.numeric(new_dat$Date[t]-new_dat$Date[1])
    
  }
  outputDays2 = append(outputDays2,days)
}
outputDays2
CL_Disp_Alives=cbind(outputDays2,CL_Disp_Alives)
colnames(CL_Disp_Alives)[1]<-"DaysElapsed"
CL_Disp_Alives$WeeksElapsed=CL_Disp_Alives$DaysElapsed/7
CL_Disp_Alives$RoundedWeek=ceiling(CL_Disp_Alives$WeeksElapsed)

CL_Disp_Alives$RoundedWeek=as.factor(CL_Disp_Alives$RoundedWeek)
CL_Disp_Alives$Species=as.factor(CL_Disp_Alives$Species)

plot(CL_Disp_Alives$CL_prox_m~CL_Disp_Alives$RoundedWeek*CL_Disp_Alives$Species)


ggplot(CL_Disp_Alives,aes(x = as.factor(RoundedWeek),y = CL_prox_m))+
  geom_boxplot(aes(fill=Species))+
  coord_cartesian(xlim=c(0,17),ylim=c(0,15000))+
  labs(x = "Rounded Week Since Release", y = "Least Cost Distance From Ramp (m)")

r1<-with(CL_Disp_Alives, tapply(CL_prox_m, RoundedWeek, mean))
plot(r1,type="l",ylab="Mean Distance From CL",xlab="Weeks Post-Release",xlim=c(0,16))
abline(v=3,col="red",lty=2)
```

Here I will look at movement speeds of Dispersal Fish
```{r}

Dispersal_Speeds=read.csv("Dispersal_Speeds.csv")
Dispersal_Speeds$MetersPerDay=Dispersal_Speeds$MetersFromLastLocation/Dispersal_Speeds$DaysElapsed
Dispersal_Speeds$WeeksElapsed=Dispersal_Speeds$DaysElapsed/7
Dispersal_Speeds$MetersPerWeek=Dispersal_Speeds$MetersFromLastLocation/Dispersal_Speeds$WeeksElapsed


Dates = make_date(year = Dispersal_Speeds$Year, month = Dispersal_Speeds$Month, day = Dispersal_Speeds$Day)
Dispersal_Speeds = cbind(Dates,Dispersal_Speeds)
colnames(Dispersal_Speeds)[1]<-"Date"    #Renaming date column to "Date"


tag_numbers = unique(Dispersal_Speeds$RewardTag1)#unique tag numbers
n_fish = length(tag_numbers)#number of fish total

outputDays2 = numeric()
for(i in 1:n_fish){
  
  new_tag = tag_numbers[i]
  new_dat = subset(Dispersal_Speeds,RewardTag1 == new_tag)
  tmax = nrow(new_dat)
  days = numeric(tmax)
  
  for(t in 2:tmax){
    
    days[t] = as.numeric(new_dat$Date[t]-new_dat$Date[1])
    
  }
  outputDays2 = append(outputDays2,days)
}
outputDays2
Dispersal_Speeds=cbind(outputDays2,Dispersal_Speeds)
colnames(Dispersal_Speeds)[1]<-"DaysPostRelease"



```
