---
title: "Lab 3 Assignment - Scale"
output: html_notebook
---
```{r, warning=F, error=F, message=F}

require(tidyverse)
require(sf)
require(AICcmodavg)
require(tigris)
require(FedData)
require(terra)
require(tidyterra)

```

## Challenge 1 (4 points)

**Build a raster with 100 rows and 100 columns. Fill the raster cells with values of a random variable drawn from a distribution of your choosing (Poisson, Normal, Uniform, etc.). Calculate the mean and variance of the values in that raster. Now increase the grain size of those cells by factors of 2, 5, and 10, combining cell values using a mean function. At each iteration, calculate the mean and variance of the values in the resulting raster. Generate 2 scatterplots that have grain size on the x-axis. Plot the mean raster value on the y-axis of the first, and variance on the y-axis of the second. What do you notice about how these values change as you "scale up" the grain size? Why do you think this pattern occurs?**

Place your code in the code chunk below so I can reproduce your analyses/figures.

```{r}
#Creating initital raster
simpRast = rast(ncol=100, nrow=100, xmin=1, xmax=6, ymin=1, ymax=6)

#Filling with values from poisson distribution
simpRast[] = rpois(ncell(simpRast), lambda=3)
plot(simpRast)

#Calculating mean and variance of cells within raster
mean(as.matrix(simpRast))
var(as.matrix(simpRast))

#Increasing grain sizes of cells by factos of 2, 5, 10, combining cells via mean function

  #factor of 2
  simpRastMean2 <- aggregate(simpRast, fact=2, fun='mean')
  mean(as.matrix(simpRastMean2))
  var(as.matrix(simpRastMean2))

  #Factor of 5
  simpRastMean5 <- aggregate(simpRast, fact=5, fun='mean')
  mean(as.matrix(simpRastMean5))
  var(as.matrix(simpRastMean5))
  
  #Factor of 10
  simpRastMean10 <- aggregate(simpRast, fact=10, fun='mean')
  mean(as.matrix(simpRastMean10))
  var(as.matrix(simpRastMean10))

  
  #Plotting mean and variance for the different grain sizes
  x=c(2,5,10)
  ymean=c(mean(as.matrix(simpRastMean2)),mean(as.matrix(simpRastMean5)),mean(as.matrix(simpRastMean10)))
  plot(ymean~x)
  
  yvar=c(var(as.matrix(simpRastMean2)),var(as.matrix(simpRastMean5)),var(as.matrix(simpRastMean10)))
  plot(yvar~x)
```

The overall mean remains the same despite the scale-up. The variance of the rasters decreases as grain size is scaled up. This is likely becuase we used the mean function to combine cells, resulting in less variance in the newly created cells. 

$\color{red}{\text{You're on the right track, although your answer is a bit hand-wavy. What I wanted to hear is that as grain size increases and we take the mean of nearby pixels, the new, larger grained pixels converge on the true mean of all the values, meaning pixels become more and more alike. +3.5}}$


## Challenge 2 (4 points)

**Identify a situation in which you might use a summary function other than the mean to calculate new cell values when you scale up the grain of a raster (e.g., median, mode, minimum, maximum, etc.). Repeat the effort from Challenge 1 using this alternate function. Again, create two scatterplots showing how the mean and variance values of the raster change as you scale up the cell size by factors of 2, 5, and 10. Do you see a similar pattern? Compare and contrast your findings with those from Challenge 1.**

*Hint: You should be able to recycle your code from Challenge 1 with only a couple of small tweaks to answer this question.*

Place your code in the code chunk below so I can reproduce your analyses/figures.

```{r}
#Creating initital raster
simpRast = rast(ncol=100, nrow=100, xmin=1, xmax=6, ymin=1, ymax=6)

#Filling with values from poisson distribution
simpRast[] = rpois(ncell(simpRast), lambda=3)
plot(simpRast)

#Calculating mean and variance of cells within raster
mean(as.matrix(simpRast))
var(as.matrix(simpRast))

#Increasing grain sizes of cells by factos of 2, 5, 10, combining cells via mean function

  #factor of 2
  simpRastMode2 <- aggregate(simpRast, fact=2, fun='modal')
  mean(as.matrix(simpRastMode2))
  var(as.matrix(simpRastMode2))

  #Factor of 5
  simpRastMode5 <- aggregate(simpRast, fact=5, fun='modal')
  mean(as.matrix(simpRastMode5))
  var(as.matrix(simpRastMode5))
  
  #Factor of 10
  simpRastMode10 <- aggregate(simpRast, fact=10, fun='modal')
  mean(as.matrix(simpRastMode10))
  var(as.matrix(simpRastMode10))

#Plotting mean and variance for the different grain sizes
  ymean2=c(mean(as.matrix(simpRastMode2)),mean(as.matrix(simpRastMode5)),mean(as.matrix(simpRastMode10)))
  plot(ymean2~x)
  
  yvar2=c(var(as.matrix(simpRastMode2)),var(as.matrix(simpRastMode5)),var(as.matrix(simpRastMode10)))
  plot(yvar2~x)
```

The mean of the raster gradually increased as I scaled up the raster which was different from challenge 1. However, variance decreased as I scaled up the raster which also happened for challenge 1. 

$\color{red}{\text{Why do you think these things are happening? And when might you use mode instead of mean to aggregate cells in a raster? +3}}$

## Challenge 3 (2 points)

**Recall that before we calculated forest cover, we cropped our NLCD raster to minimize its size and the computing effort necessary from our poor little computers. How might that affect our ability to evaluate the scale at which five-lined skinks respond to forest cover? Why?**

This cropping may inhibit our ability to identify the scale at which five-lined skins respond to forest cover if the true effect/scale takes place at a scale larger than the area we cropped to. This is similar to issues brought to light by the paper we talked about in discussion this past week. Many studies found that the greatest effect was at the smallest or largest scale evaluated. Therefore, it could be an issue in this specific example if we were to find the greatest effect at the largest scale we could evaluate because our crop wont let us evaluate a larger scale which is where the true effect might be.

$\color{red}{\text{Perfect. +2}}$

## Challenge 4 (4 points)

**In the lab, we measured forest cover at 1 km and 5 km. Extract forest cover proportions around each sample point for 100 m, 500 m, 1 km, 2 km, 3 km, 4 km, and 5 km scales. Examine the correlation between these 7 variables (remember the chart.Correlation() function). What patterns do you notice in correlation among these variables?**

*Hint: Recall the for loop we used to calculate this variable at two scales... could you make a small addition here to look at more scales?*

```{r}

sites = st_read("/vsicurl/https://github.com/ValenteJJ/SpatialEcology/raw/main/Week3/reptiledata.shp") %>% 
  filter(management!='Corn')
st_crs(sites) = "+proj=aea +lat_0=23 +lon_0=-96 +lat_1=29.5 +lat_2=45.5 +x_0=0 +y_0=0 +datum=NAD83 +units=m +no_defs"
head(sites)


states = states() %>% 
  filter(NAME %in% c('Alabama', 'Florida', 'Georgia')) %>% 
  st_transform(crs(sites, proj=T))



ggplot()+
  geom_sf(data = states)+
  geom_sf(data = sites)

presAbs = read.csv('https://raw.githubusercontent.com/ValenteJJ/SpatialEcology/main/Week3/reptiles_flsk.csv')

sites = sites %>% 
  left_join(presAbs, by='site')

#Extract x and y coordinates of the bounding box
studyArea = st_bbox(sites) + c(-10000, -10000, 10000, 10000)
studyArea = st_as_sfc(studyArea)


ggplot()+
  geom_sf(data = states)+
  geom_sf(data = studyArea, fill=NA, color='red')+
  geom_sf(data = sites)

nlcd = get_nlcd(studyArea,
                label='studyArea',
                year = 2016,
                dataset = 'landcover',
                landmass = 'L48'
)



plot(nlcd, 1, legend=T, plg=list(cex=0.5))
plot(st_geometry(sites), add=T, pch=16)


forest = nlcd %>% 
  setValues(0)

forest[nlcd=='Deciduous Forest' | nlcd=='Evergreen Forest' | nlcd=='Mixed Forest'] = 1
plot(forest)
plot(st_geometry(sites), add=T, pch=16, col='black')


#######Buffer Function
bufferCover = function(shp, size, landcover){
  buffArea = (pi*size^2)/10000
  grainArea = (prod(res(landcover)))/10000
  
  buffi = st_buffer(shp[i,], dist=size)
  cropi = crop(landcover, buffi, mask=T)
  numCells = global(cropi, 'sum', na.rm=T)
  forestHa = numCells * grainArea
  propForest = forestHa / buffArea
  
  return(propForest)
}

#####For-loop
for100m = as.vector(rep(NA, nrow(sites)))
for500m = as.vector(rep(NA, nrow(sites)))
for1km = as.vector(rep(NA, nrow(sites)))
for2km = as.vector(rep(NA, nrow(sites)))
for3km = as.vector(rep(NA, nrow(sites)))
for4km = as.vector(rep(NA, nrow(sites)))
for5km = as.vector(rep(NA, nrow(sites)))

for(i in 1:nrow(sites)){
  for100m[i] = bufferCover(sites, 100, forest)
  for500m[i] = bufferCover(sites, 500, forest)
  for1km[i] = bufferCover(sites, 1000, forest)
  for2km[i] = bufferCover(sites, 2000, forest)
  for3km[i] = bufferCover(sites, 3000, forest)
  for4km[i] = bufferCover(sites, 4000, forest)
  for5km[i] = bufferCover(sites, 5000, forest)
}

#Mutating results into a new dataframe
forestData = sites %>% 
  mutate(for100m = unlist(for100m),
         for500m = unlist(for500m),
         for1km = unlist(for1km),
         for2km = unlist(for2km),
         for3km = unlist(for3km),
         for4km = unlist(for4km),
         for5km = unlist(for5km))

head(forestData)

#Creating coorelation between the 7 buffer lengths
forestData %>% 
  as.data.frame() %>% 
  select(coords_x1, for100m, for500m, for1km, for2km, for3km, for4km, for5km) %>% 
  PerformanceAnalytics::chart.Correlation(histogram=F)

```
I notice that adjacent scales (Ex. 1km and 2km, or 4km and 5km) are highly coorelated with one another and are more linear. Smaller scales tended to have more variability.

$\color{red}{\text{And scales that are further apart have less correlation than those close together. +3}}$

## Challenge 5 (4 points)

**Fit 8 logistic regression models (a null model and one for each of the 7 forest scales). Compare these models using AICc. Which scale do you think represents the critical or characteristic scale at which forest cover affects skink presence? Is this scale clearly better than the others, or is there some ambiguity? What are some mechanisms by which forest cover could affect skink presence at this scale? What is your overall conclusion regarding how forest cover affects skink presence (i.e., take a look at the betas)?**

Place your R code in the chunk below.
```{r}
modelNull = glm(pres~1, family='binomial', data=forestData)
model100m = glm(pres~for100m, family='binomial', data=forestData)
model500m = glm(pres~for500m, family='binomial', data=forestData)
model1km = glm(pres~for1km, family='binomial', data=forestData)
model2km = glm(pres~for2km, family='binomial', data=forestData)
model3km = glm(pres~for3km, family='binomial', data=forestData)
model4km = glm(pres~for4km, family='binomial', data=forestData)
model5km = glm(pres~for5km, family='binomial', data=forestData)
# 
# install.packages("pbapply",dependencies=TRUE)
# library(AICcmodavg)
aictab(list(modelNull, model100m, model500m, model1km, model2km, model3km, model4km, model5km), 
       modnames=c('Null', '100 m','500 m','1 km','2 km', '3 km', '4 km','5 km'))

effects = data.frame(model = c('100m', '500m', '1km', '2km', '3km','4km','5km'),
           beta = c(summary(model100m)$coefficients[2,1],summary(model500m)$coefficients[2,1],
                    summary(model1km)$coefficients[2,1], summary(model2km)$coefficients[2,1],
                    summary(model3km)$coefficients[2,1], summary(model4km)$coefficients[2,1],
                    summary(model5km)$coefficients[2,1]),
           se = c(summary(model100m)$coefficients[2,2],summary(model500m)$coefficients[2,2],
                  summary(model1km)$coefficients[2,2], summary(model2km)$coefficients[2,2],
                  summary(model3km)$coefficients[2,2],summary(model4km)$coefficients[2,2],
                  summary(model5km)$coefficients[2,2]))

effects = effects %>% 
  mutate(lcl = beta - 1.96*se,
         ucl = beta + 1.96*se)
```

According to AICc alone, it appears that the 2, 3, 4, and 5km models are all competing models that would explain the data the best. I personally feel like there is some ambiguity to his based on AICc alone. Perhaps we didn't examine the proper scale for this relationship? According to the betas there does appear to be a relationship between forest cover and skink presence. The beta is clearly higher than the others starting at 2km and stays roughly the same for all other tested scales. It appears that skink presence is heavily effected by forest cover starting a 2km scale and is not more effected at scales over 2km. 

$\color{red}{\text{What about the mechanisms operating at these broader scales? +3}}$

## Challenge 6 (2 points)

**If you encounter ambiguity in identifying the characteristic scale of an effect, can you come up with a clever way to condense the information in the multi-scale variables into just one or two? When might it be ok to include two covariates in the same model (think multiple regression) that represent the same ecological feature measured at different scales (e.g., forest cover at 1 km AND forest cover at 5 km in the same model)? I can think of both a biological and a statistical answer to this question.**

No idea.

$\color{red}{\text{Lots of answers fit in here, but I can't give you credit for this one. You could find a biological justification for different scales at which the skinks might respond to forest cover and focus on those. You could combine variables using a PCA, you could simply select some that are highly correlated with the others as "representative."  You also want to be careful to not include two variables in the same model if they are highly correlated with one another, as this causes variance inflation problems. +0}}$
