<<<<<<< HEAD
---
title: "R Notebook"
output: html_notebook
---

```{r}
require(terra)
require(tidyverse)
require(pgirmess)
require(ncf)
require(geoR)
```


# Background

Researchers set up a square plot with a dimension of 130 m on each side. Within the plot they established blue vane traps in a 5 m grid pattern to sample bee abundance. Every day for 10 days the researchers emptied the blue vane traps and recorded the number of bees collected. After the 10-day sampling period, they calculated the mean daily number of bees captured in the trap each day.

# Challenge 1 (2 points)

Import the blueVanes.csv file from GitHub. Plot the mean bee abundance as a function of the x and y coordinates. Upon initial inspection, does there appear to be spatial dependence in the bee response data?

```{r}
vaneMatrix = read.csv("https://github.com/ValenteJJ/SpatialEcology/raw/main/Week6/blueVanes.csv")#had to change /blob/main to /raw/main
head(vaneMatrix)

#plotting 
ggplot(vaneMatrix, aes(x=x, y=y, color=meanBee))+
  geom_point()+
  theme_bw()+
  theme(panel.grid=element_blank())+
  scale_color_gradient(low='blue', high='red')
```

Yes, it does appear that there might be some spatial dependence in the bee data. 

# Challenge 2 (3 points)

Empirically test for evidence of spatial dependence in bee abundance by first building a corelogram, then building a semi-variogram. Be sure to provide an envelope for assessing whether the values in each figure are statistically different from what would be expected if the bee values were randomly distributed. Describe the patterns you see, and your interpretation with regards to the spatial dependence in mean bee abundance. Be sure to identify the maximum distance at which you should interpret your figures (2/3 of the maximum distance between points).

```{r}
#Correlogram
  #Distances among all points
  coords = cbind(vaneMatrix$x, vaneMatrix$y)
  colnames(coords) = c('x', 'y')
  distMat = as.matrix(dist(coords))
  
  #Setting max distance 
  maxDist = (2/3) * max(distMat)

  bee_correlog = pgirmess::correlog(coords, vaneMatrix$meanBee, method='Moran', nbclass=14, alternative = 'two.sided')
  round(bee_correlog, 2)
  
  #Plotting results using ggplot()
  tmp = data.frame(bee_correlog) %>%
  mutate(significant = ifelse(p.value < 0.05, 'yes', 'no'))
  
ggplot(tmp, aes(x=dist.class, y=coef))+
  geom_line()+
  geom_point(aes(color=significant, size=n))+
  scale_color_manual(values=c('black', 'red'))+
  theme_bw()+
  theme(panel.grid=element_blank())+
  geom_hline(yintercept=0, linetype='dashed')+
  geom_vline(xintercept=maxDist, color='red')+
  scale_x_continuous(breaks=seq(0, 130, 5))

#Semi-Variogram
  #Creating geoR object
  geoBee = as.geodata(vaneMatrix)
  plot(geoBee)
  
  #Building the variogram
  empGeor = variog(geoBee, max.dist=maxDist,
                   breaks=seq(0, maxDist, by=3))#should I change the "by" argument?
  plot(empGeor)
  
  #CREATING ENVELOPE for VARIOGRAM
  empEnv = variog.mc.env(geoBee, obj.var=empGeor, nsim=99)
  plot(empGeor, envelope=empEnv)
```
The maximum distance we should interpret our figures is 2/3 of the maximum distance between points. This would be right around 90 m. According to the correlogram we had significant spatial dependence for all distance classes less than 90 m except 2. The variogram, however, showed significant spatial dependence at a lesser rate, with the groups deemed significant being roughly 20 and less and 80 and higher. However, no matter how you look at it, both graphs did show signs of spatial dependence in mean bee abundance for our data. 


# Challenge 3 (4 points)

Fit a simple linear model (assuming a normally distributed error term) that models mean bee abundance as a function of an intercept only (no other covariates in the model). Compare the model-estimated intercept to the bee abundance values averaged across all sites. What do you notice?  

Empirically test for evidence of spatial autocorrelation IN THE MODEL RESIDUALS by first building a correlogram, then building a semi-variogram. Be sure to provide an envelope for assessing whether the values in each figure are statistically different from what would be expected if the bee values were randomly distributed in space. Compare these figures to those built for the raw data. What similarities or differences to you notice?

```{r}
#Fitting simple linear model of mean bee abundance as a function on an intercept
interceptfit=lm(vaneMatrix$meanBee~1)
mean(vaneMatrix$meanBee)#They are the same

#Testing for autocorrelation in the RESIDUALS (x,y of bees, z is resids)
  
  #Creating resids column in original data
  vaneMatrix$resids=resid(interceptfit)
  
  #Correlogram
  #Distances among all points
      bee_correlog_resids = pgirmess::correlog(coords, vaneMatrix$resids, method='Moran', nbclass=14, alternative =
                                          'two.sided')
      round(bee_correlog, 2)
  
  #Plotting results using ggplot()
  tmp2 = data.frame(bee_correlog_resids) %>%
  mutate(significant = ifelse(p.value < 0.05, 'yes', 'no'))
  
ggplot(tmp2, aes(x=dist.class, y=coef))+
  geom_line()+
  geom_point(aes(color=significant, size=n))+
  scale_color_manual(values=c('black', 'red'))+
  theme_bw()+
  theme(panel.grid=element_blank())+
  geom_hline(yintercept=0, linetype='dashed')+
  geom_vline(xintercept=maxDist, color='red')+
  scale_x_continuous(breaks=seq(0, 130, 5))

  #Semi-variogram with envelope

  #Creating geoR object
  geoBee_resids = as.geodata(vaneMatrix)
  
  #Building the variogram
  empGeor_resids = variog(geoBee, max.dist=maxDist,
                   breaks=seq(0, maxDist, by=3))#should I change the "by" argument?
  plot(empGeor_resids)
  
  #CREATING ENVELOPE for VARIOGRAM
  empEnv_resids = variog.mc.env(geoBee_resids, obj.var=empGeor_resids, nsim=99)
  plot(empGeor_resids, envelope=empEnv_resids)

  #Compare figures to those built for the raw data (Challenge 2)
  #They appear to be the same
```
The model estimated intercept was identical to the bee abundance averaged across all sites. These figures (the newly created correlogram and variogram) look identical to the correlogram and variogram produced for the raw data. The only thing that looks a little different is the envelopes for the variograms. 




# Challenge 4 (4 points)

Conventional wisdom indicates that the distribution of floral resources may substantially influence bee abundance. In addition, I am  completely making up the following fact: the average bee has a perceptual range of about 15 m. Below I have provided code for importing a raster that shows the number of flowers counted in 1 m grid cells across the entire study area. Use the focal() function from the previous lab to generate a new raster in which each cell value represents the mean number of flowers in cells within a 15 m radius. Plot the distribution of the blue vane traps over the 15 m focal raster. No text required here, just show me the plot.

```{r}
floral = rast('https://github.com/ValenteJJ/SpatialEcology/raw/main/Week6/floralResources.tif')

#Bee has perceptual range of 15 m, flowers is 1 m grid cells
#focal() to generate raster where each cell is mean number of flowers in cells within a 15 m radius

meanFlower15m=focal(floral,w=15,fun="mean")
plot(meanFlower15m)
points(x = vaneMatrix$x,y=vaneMatrix$y)
```

# Challenge 5 (4 points)

Extract the focal flower value from the raster you just created at each point using the extract() function. You now have an estimate for the average number of flowers per meter around each blue vane trap. Fit a second linear model, but this time include average floral abundance within 15 m as a covariate. Export the residuals from this fitted model, and again, examine a corelogram and variogram of the residuals. As always, be sure to plot random envelopes. Now what do you notice?

```{r}
#Using extract function
focalMean = terra::extract(meanFlower15m,coords)
focalMean=as.vector(focalMean)
dat=cbind(focalMean,vaneMatrix)
str(dat)

#Fitting Linear Model with floral abundance within 15m as a covariate
results=lm(dat$meanBee~dat$focal_mean)
summary(results)

#Testing for autocorrelation in the RESIDUALS (x,y of bees, z is resids)
  
  #Creating resids column in original data
  dat$resids2=resid(results)
  
  #Correlogram
  #Distances among all points
      bee_correlog_resids2 = pgirmess::correlog(coords, dat$resids2, method='Moran', nbclass=14, alternative =
                                          'two.sided')
      round(bee_correlog_resids2, 2)
  
  #Plotting results using ggplot()
  tmp3 = data.frame(bee_correlog_resids2) %>%
  mutate(significant = ifelse(p.value < 0.05, 'yes', 'no'))
  
ggplot(tmp3, aes(x=dist.class, y=coef))+
  geom_line()+
  geom_point(aes(color=significant, size=n))+
  scale_color_manual(values=c('black', 'red'))+
  theme_bw()+
  theme(panel.grid=element_blank())+
  geom_hline(yintercept=0, linetype='dashed')+
  geom_vline(xintercept=maxDist, color='red')+
  scale_x_continuous(breaks=seq(0, 130, 5))

  #Semi-variogram with envelope

  #Creating geoR object
  geoBee_resids2 = as.geodata(dat)
  
  #Building the variogram
  empGeor_resids2 = variog(geoBee_resids2, max.dist=maxDist,
                   breaks=seq(0, maxDist, by=3))#should I change the "by" argument?
  plot(empGeor_resids2)
  
  #CREATING ENVELOPE for VARIOGRAM
  empEnv_resids2 = variog.mc.env(geoBee_resids2, obj.var=empGeor_resids2, nsim=99)
  plot(empGeor_resids2, envelope=empEnv_resids2)
```

I notice that there isn't really any evidence for spatial dependency if you take into account the distribution of flowers on the landscape. The bees appear to be randomly distributed among the flowers. 

# Challenge 6 (3 points)

What is your assessment regarding the distribution of bee abundance in this study area? Does it exhibit spatial dependence, and if so, what is driving that pattern?

So, it does appear (without taking anything into account) that there is spatial dependence among the bees in the landscape. However, in challenge 5, we took into account the distribution of flowers in the landscape. Upon doing so, we found out that the distribution of bees (when taking flowers into account) is not significantly different from random, thus there is no spatial dependence. Flowers appear to be driving the bee distribution pattern and this can be accounted for as we showed in challenge 5. 
