---
title: "Marcus Prull Lab 2"
output: html_notebook
---

#Loading necessary packages
```{r}
require(tidyverse)
require(sf)
require(tigris)
require(geodata)
require(terra)
require(colorspace)
require(tidyterra)
require(ggnewscale)
require(cowplot)
require(FedData)
```

#Loading in necessary study area shapefile
```{r}
Neely = st_read("/vsicurl/https://github.com/PrullMarcus/SpatialEcology/raw/main/Lab%202/NeelyData/FINAL.shp") 
```

#Creating bounding box around study area
```{r}

studyArea = st_bbox(Neely)
studyArea

studyArea = studyArea + c(-2000, -2000, 2000, 2000)
studyArea

studyArea = st_as_sfc(studyArea)

plot(studyArea)
```

#Obtaining state of Alabama shapefile for zoomed out map
```{r}
alabama = states() %>% 
  filter(NAME=='Alabama')

plot(st_geometry(alabama))

st_crs(alabama)$proj4string
```

#Transforming the Alabama coordinate system to match that of my study area shapefile
```{r}
alabama = st_transform(alabama, st_crs(studyArea))
```

#Plot zoomed out map of study area 
```{r}
plot(st_geometry(alabama))
plot(studyArea, col='red', add=T)
```

#Zoomed in map of the reservoir
```{r}
plot(studyArea)
plot(Neely,add=T)
```

#Obtaining landcover data for the area surrounding Neely Henry
```{r}
NeelyLC=get_nlcd(template=studyArea,year=2019,
                 dataset="landcover",
                 label="Neely Henry",
                 landmass="L48")
```

#Transforming shapefile of Neely to match the LC raster
```{r}
Neely = st_transform(Neely, st_crs(NeelyLC))
```

#Plotting my shapefile in addition to the landcover data.
```{r}

plot(NeelyLC)
plot(Neely, add=T)

```

#Creating zoomed out map
```{r}
zoomedOut = ggplot()+
  geom_sf(data = alabama)+
  geom_sf(data = studyArea, fill=NA, color='red')

zoomedOut
```

#Cleaning up zoomed out map
```{r}
zoomedOut = zoomedOut +
  theme_bw()+ #Gets rid of most of the colors
  theme(panel.grid=element_blank(), #Removes the gridlines
        axis.text=element_blank(), #Removes the axis text
        axis.ticks=element_blank(), #Removes the axis tick marks
        panel.border=element_blank()) #Removes the frame around the graph

zoomedOut
```
```{r}

zoomedIn = ggplot()+
  geom_spatraster(data = NeelyLC)

zoomedIn
```

#Starting to make the nice zoomed in map
```{r}

zoomedIn = ggplot()+
  geom_spatraster(data = NeelyLC)+
  geom_sf(data = Neely, fill="blue", color='black')

zoomedIn
```

#Cleaning up the zoomedIn map
```{r}
zoomedIn = zoomedIn+
  new_scale_fill()+
  theme_bw()+
  theme(panel.grid=element_blank(),
        axis.text.x=element_text(angle=90, vjust=0.5, size=6))

zoomedIn
  
```
#Using cowplot to put my two maps together
```{r}
plot_grid(zoomedIn, zoomedOut)
```
#Taking out legend
```{r}
legend = get_legend(zoomedIn)
ggdraw(legend)
```
#Creating an object for the legened
```{r}
legend = get_legend(zoomedIn+theme(legend.direction='vertical'))
ggdraw(legend)

zoomedIn = zoomedIn + theme(legend.position='none')
```
#Adding the legend and zoomedOut maps to the final map
```{r}
supportingFig = plot_grid(ggdraw(legend), zoomedOut, ncol=2, rel_heights = c(1, 1))

supportingFig
```
#Adding my zoomedIn map to the legend and zoomedOut map
```{r}
plot_grid(zoomedIn, supportingFig, rel_widths=c(0.7, 1))
```
#Neely Henry (left panel) is a roughly 15,000 acre reservoir located on the Coosa River. I am tracking two difference species of fish on this reservoir over the course of two years. I chose to include a shapefile I created of the lake and landcover use of the surrounding area in my final map. It is interesting to see how urbanized the North end of the lake is near Gadsden, Alabama, compared to the southern part of the lake which is more heavily forested. Additionally, the right panel shows Neely Henry's location relative to the rest of the state of Alabama.




